---
title: "Shine bright like an open source app"
subititle: "Basics"
author: "Ottavia M. Epifania"
format: 
  revealjs: 
    page-layout: custom
    theme: simple
server: shiny
code-annotations: hover
---

# The very first app 

## 

`appExamples/basic-App.R`

::: {.panel-tabset}

## The app

```{r}
#| panel: center
selectInput(inputId = "mySelection",  
            label = h2("Select box"), 
    choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3), 
    selected = 1) 
verbatimTextOutput("myOutput")
```


```{r}
#| context: server
output$myOutput = renderPrint({
    paste(paste("This is my choice"), input$mySelection)
})
```


## The code 

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|1-7|8-12|13"
ui = fluidPage(selectInput(inputId = "mySelection",  # <1>
                          label = h3("Select box"),  # <2>
                          choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3), # <3>
                          selected = 1), # <4>
         verbatimTextOutput( outputId = "myOutput") # <5>
         )
server = function(input, output) {
           output$myOutput = renderPrint({ # <6>
             paste(paste("This is my choice"),  input$mySelection) # <7>
           })
}
shinyApp(ui, server)
```

1. Define the ID of the input in the UI (internal use)
2. Define the label of the element (external use)
3. Options (labels) and their internal labels
4. Default selected option
5. Define the label (internal use) of the outuput in the UI (where the results elaborated by the server will be stored)
6. Function in the server to render the results (consistent with the UI definition)
7. Call for the label of the input in the UI

:::

## App layout

```{r eval=FALSE}
#| echo: true
#| code-line-numbers: "|1|3-6|8-11"
fluidPage()   # <1> 

fluidRow(      #  <2>
  column(6, ), # <2> 
  column(6, )  #  <2> 
)   # <2>

sidebarLayout(    # <3> 
sidebarPanel(...), # <3>  
mainPanel(...)    # <3> 
) # <3>

# and many more!
```

1. Rows with columns within 12-unit wide grid
2. Custom the page in two columns of width 6 out of the 12-unit on a single row
3. Automatic layout with sidebar and main area

[Complete list](https://shiny.rstudio.com/gallery/) (look for "Application layout")


## The app

```{r}
#| panel: sidebar
selectInput(inputId = "dataset", # name of the input (for the server)
                  label = "Choose a dataset:", # name of the input (for the users)
                  choices = c("rock", "pressure", "cars")) # options (for  both 
                                                          # users & server)
```


```{r out.width="30%"}
#| panel: fill
 plotOutput("graph" )
verbatimTextOutput("summary")

```





```{r}
#| context: server

output$graph <- renderPlot({
    if(input$dataset == "rock"){ # call the input and its options with their label
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    plot(data[, c(1:2)])
  })

output$summary <- renderPrint({
    if(input$dataset == "rock"){
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    summary(data[, c(1:2)])
  })

```




## {.smaller}

`appExamples/first-app.R`

::: {.panel-tabset}

## UI

```{r}
#| echo: true
#| code-line-numbers: "|3,7|8,11|4-6|9|10"
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(    
      selectInput(inputId = "dataset",  
                  label = "Choose a dataset:", 
                  choices = c("rock", "pressure", "cars")) 
    ),
    mainPanel(
      plotOutput( "graph"  ), # <1> 
      verbatimTextOutput( "summary") # <2>
    ) 
  )
)
```

1. Define that specific type of output labeled "graph" is a graph
2. Define that specific type of output labeled "summary" is a code 

## Server

```{r}
#| echo: true
#| code-line-numbers: "|2,8|9,15|3-7|10-14|7,14"
server <- function(input, output){
  output$graph <- renderPlot({  # <1>
   data =  switch(input$dataset,
           rock = rock,
           pressure = pressure,
           cars = cars)
    plot(data[, c(1:2)])
  })
output$summary <- renderPrint({ # <2>
   data =  switch(input$dataset,
           rock = rock,
           pressure = pressure,
           cars = cars)
    summary(data[, c(1:2)])
  })
}
```


1. renderPlot is the function in the server mirroring the plotOuput funtcion in the UI
2. renderPrint is the function in the server mirroring the verbatimTextOutput in the UI

## Together

```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::


# Don't repeat yourself

## The "shelf" solution {.smaller}

::: {.panel-tabset}




## UI

```{r}
#| echo: true
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "dataset",
                  label = "Choose a dataset:", 
                  choices = c("rock", "pressure", "cars")) 
    ),
    mainPanel(
      plotOutput("graph"),
      verbatimTextOutput("summary")
    ) 
  )
)
```



## Server

```{r}
#| echo: true
#| code-line-numbers: "|2-8|9-11|13,16"
server <- function(input, output){ 
values <- reactiveValues()  # <1> 
  dataInput <- reactive({   # <2>
     data =  switch(input$dataset, # <2>
           rock = rock, # <2>
           pressure = pressure, # <2>
           cars = cars) # <2>
  }) # <2>
observe({            # <3>          
    values$data <- data.frame(dataInput())  # <3>
  })                                       # <3>
output$graph <- renderPlot({
       plot(values$data[, c(1:2)])  # <4>
    })
output$summary <- renderPrint({
    summary(values$data[, c(1:2)]) # <4>
})
}

```

1. Define a variable as reactiveValues (i.e., a variable that reacts to what happen in the setting configuration of the UI)
2. Define the data as reacting to what happens in the UI
3. Link the reactive variable with the data input (i.e., create the shelf)
4. Use the shelf for the analysis

## Together

```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::

## `reactive`()` {.smaller}


::: {.panel-tabset}

## UI 

```{r}
#| echo: true
#| code-line-numbers: "|8|"
ui = fluidPage(
    sidebarLayout(sidebarPanel(
        selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices  = c("rock", "pressure", "cars"))
      ),
mainPanel(
      plotOutput("graph"), 
      verbatimTextOutput( 
          "summary"   
      ))))
```


## Server 

```{r}
#| echo: true
#| code-line-numbers: "|2-7|9,13"
server <- function(input, output){
    dataInput <- reactive({ # <1>
    data =  switch(input$dataset, # <1> 
           rock = rock,  # <1>
           pressure = pressure, # <1> 
           cars = cars) # <1>
    }) # <1>
output$graph <- renderPlot({
      data <- dataInput() # <2>
      plot(data[, c(1:2)])
    })
output$summary <- renderPrint({
      data <- dataInput() # <2>
      summary(data[, c(1:2)])
    })
  }
```

1. Define the reactive dataset with the reactive function
2. Use the reactive dataset for the analysis

## Together


```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::

# Prevent the action

## Action buttons

They allow for conditioning a specific action to an event

As such, the app will show the output only if a specific event is observed 

Both shelf and reactive solutions work with the action buttons

We are studying the solution with the reactive functions because they are more common and used (more documentation is availble)



