---
title: "Slow and steady wins the race - `eventReactive()`"
author: "Ottavia M. Epifania"
format: 
  revealjs: 
    page-layout: custom
    theme: simple
    slide-number:  c/t
    show-slide-number: all
server: shiny
code-annotations: hover
---


# `eventReactive()` and `actionButton()`

## 

These functions are usually employed together: 

:::{.callout-note}
## The action button

```{r}
#| echo: true
actionButton("start", 
             "Press here to do something")
```

:::

:::{.callout-note}

## The action button controlling the actions

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|3-4|7-9|"
ui = fluidPage(
  [...]
  actionButton("start",  # <1> 
             "Press here to do something") #<1>
)
server = server(input, output) {
  data = eventReactive(input$start, # <2>
                    [...]   # <3>
                       )
}
```

1. Define the button label to be used in the server 
2. Define the action after which the button event becomes reactive
3. Code and operations to be performed after the button has been pressed
:::

## An example

```{r}
#| panel: sidebar
selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices = list("rock" = 1, 
                                   "pressure" = 2,
                                   "cars" = 3))
actionButton("load", "Select dataset")
```

```{r}
#| panel: fill
plotOutput("graph") 
verbatimTextOutput("summary")
```

```{r}
#| context: server
dataInput = eventReactive(input$load, {
      if(input$dataset == 1){
        data <- rock
      } else if (input$dataset == 2 ){
        data <- pressure
      } else if (input$dataset == 3) {
        data <- cars
      } 
    })

output$graph <- renderPlot({
      data = dataInput()
      plot(data[, c(1:2)])
    })
    
output$summary <- renderPrint({
      data = dataInput()
      summary(data[, c(1:2)])
    })
```

## {.smaller}

::: {.panel-tabset}

## UI 

```{r}
#| echo: true
#| code-line-numbers: "|8|"
ui <- fluidPage(
    sidebarLayout(sidebarPanel(
        selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices = list("rock" = 1, 
                                   "pressure" = 2,
                                   "cars" = 3)), 
      actionButton("load", "Select dataset") # <1>
      ),
mainPanel(
      plotOutput(
          "graph"   
        ), 
      verbatimTextOutput( 
          "summary"   
      ))))
```

1. Action button visible in the UI

## Server 

```{r}
#| echo: true
#| code-line-numbers: "|2-11|13,17"
server <- function(input, output){
    dataInput <- eventReactive(input$load, { # <1> 
      if(input$dataset == 1){                # <2>
        data <- rock          # <2>
      } else if (input$dataset == 2 ){ # <2>
        data <- pressure # <2>
      } else if (input$dataset == 3) { # <2>
        data <- cars # <2>
      }  # <2>
    }  # <2>
    )# <2>
output$graph <- renderPlot({
      data <- dataInput()  # <3>
      plot(data[, c(1:2)])
    })
output$summary <- renderPrint({
      data <- dataInput()  # <3>
      summary(data[, c(1:2)])
    })
  }
```

1. Use eventReactive to make the app "react" after a specific action (e.g., clicking on the action button)
2. Normal code, in this case defining the specific dataset 
3. Actually makes the dataset usable for the analysis

## Together


```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::
