---
title: "Slow and steady wins the race - `eventReactive()`"
author: "Ottavia M. Epifania"
format: 
  revealjs: 
    page-layout: custom
    theme: serif
server: shiny
code-annotations: hover
---


# `eventReactive()` and `actionButton()`

## 

These functions are usually employed together: 

:::{.callout-note}
## The action button

```{r}
#| echo: true
actionButton("start", 
             "Press here to do something")
```

:::

:::{.callout-note}

## The action button controlling the actions

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|3-4|7-9|"
ui = fluidPage(
  [...]
  actionButton("start",  # <1> 
             "Press here to do something") #<1>
)
server = server(input, output) {
  data = eventReactive(input$start, # <2>
                    [...]   # <3>
                       )
}
```

1. Define the button label to be used in the server 
2. Define the action after which the button event becomes reactive
3. Code and operations to be performed after the button has been pressed
:::

## An example

```{r}
#| panel: sidebar
selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices = list("rock" = 1, 
                                   "pressure" = 2,
                                   "cars" = 3))
actionButton("load", "Select dataset")
```

```{r}
#| panel: fill
plotOutput("graph") 
verbatimTextOutput("summary")
```

```{r}
#| context: server
dataInput = eventReactive(input$load, {
      if(input$dataset == 1){
        data <- rock
      } else if (input$dataset == 2 ){
        data <- pressure
      } else if (input$dataset == 3) {
        data <- cars
      } 
    })

output$graph <- renderPlot({
      data = dataInput()
      plot(data[, c(1:2)])
    })
    
output$summary <- renderPrint({
      data = dataInput()
      summary(data[, c(1:2)])
    })
```

## The code 


::: {.panel-tabset}

## UI 

```{r}
#| echo: true
#| code-line-numbers: "|8|"
ui <- fluidPage(
    sidebarLayout(sidebarPanel(
        selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices = list("rock" = 1, 
                                   "pressure" = 2,
                                   "cars" = 3)), 
      actionButton("load", "Select dataset")
      ),
mainPanel(
      plotOutput(
          "graph"   
        ), 
      verbatimTextOutput( 
          "summary"   
      ))))
```


## Server 

```{r}
#| echo: true
#| code-line-numbers: "|2-11|13,17"
server <- function(input, output){
    dataInput <- eventReactive(input$load, {
      if(input$dataset == 1){
        data <- rock
      } else if (input$dataset == 2 ){
        data <- pressure
      } else if (input$dataset == 3) {
        data <- cars
      } 
    } 
    )
output$graph <- renderPlot({
      data <- dataInput()
      plot(data[, c(1:2)])
    })
output$summary <- renderPrint({
      data <- dataInput()
      summary(data[, c(1:2)])
    })
  }
```


## Together


```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::
