---
title: "Download Data and graphs"
author: "Ottavia M. Epifania"
format: 
  revealjs: 
    page-layout: custom
    theme: serif
server: shiny
code-annotations: hover
---

#


## 

The specific action button for the download

```{r}
#| echo: true
downloadButton("download", "Download this thing")
```

The button can be used to download almost anything. By default (if nothing is specified through the download handler, it downloads the source file)

The download action activated by the button is handled by a download handler

```{r}
#| echo: true
#| eval: false
downloadHandler("filename", file)
```


## An example

```{r}
#| panel: sidebar
selectInput(inputId = "dataset",
                    label = "Choose a dataset:",
                    choices = list("rock" = 1, 
                                   "pressure" = 2,
                                   "cars" = 3))
actionButton("load", "Select dataset")
```

```{r}
#| panel: fill
plotOutput("graph") 
downloadButton("downloadPlot", "Download plot")
verbatimTextOutput("summary")
downloadButton("downloadSummary", "Dowload summary")
```

```{r}
#| context: server

library(tidyverse)
dataInput = eventReactive(input$load, {
      if(input$dataset == 1){
        data <- rock
      } else if (input$dataset == 2 ){
        data <- pressure
      } else if (input$dataset == 3) {
        data <- cars
      } 
    })
output$graph <- renderPlot({
      df = dataInput()
      plot(df[, c(1:2)])
    })
ddata <- reactive({
  df = dataInput()
  descript_data= df %>%
    summarise( m_1 = mean(df[,1]), sd_1 = sd(df[,1]),
      m_2 = mean(df[,2]), sd_2 = sd(df[,2])
    )
  colnames(descript_data) <- gsub("1", colnames(df)[1], colnames(descript_data))
  colnames(descript_data) <- gsub("2", colnames(df)[2], colnames(descript_data))
 descript_data
})  
output$summary <- renderPrint({
 ddata()
  })
output$downloadSummary <- downloadHandler(
    filename = function() {
      paste("descriptive-",input$dataset, ".csv", sep = "")
    },
    content = function(file) {
      write.table(ddata(), file, sep = ";",
                  row.names = FALSE)
    }
  )
  output$downloadPlot <- downloadHandler(
    
    filename = function() {
      paste0("plot-", input$dataset, ".png")
    },
    
    content = function(file) {
      
      png(file, width = 800, height = 600)
      
      df <- dataInput()
      plot(df[, 1:2])
      
      dev.off()
    }
  )
```




## {.smaller}

::: {.panel-tabset}

## UI 

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|12,14|"
ui = fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "dataset",
                  label = "Choose a dataset:",
                  choices = c("rock", 
                                 "pressure",
                                 "cars")),
      actionButton("load", "Select dataset"),
    ), 
    mainPanel(plotOutput("graph"),
              downloadButton("downloadPlot", "Download plot"), # <1>
              verbatimTextOutput("summary"), 
              downloadButton("downloadSummary", "Dowload summary")) # <2>
  )
)
```

1. Dowload button for the plot
2. Download button for the dataset

## Server

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|12-20|24-32|33-43|"
server = function(input, output){
 dataInput = eventReactive(input$load, {
    switch(input$dataset,
           rock = rock,
           pressure = pressure,
           cars = cars)
  })
  output$graph <- renderPlot({
    df = dataInput()
    plot(df[, c(1:2)])
  })
ddata = reactive({  # <1>
    df = dataInput()  # <1>
    descript_data= df %>% # <1>
      summarise( m_1 = mean(df[,1]), sd_1 = sd(df[,1]),# <1>
                 m_2 = mean(df[,2]), sd_2 = sd(df[,2])) # <1>
    colnames(descript_data) <- gsub("1", colnames(df)[1],  colnames(descript_data)) # <1>
    colnames(descript_data) = gsub("2", colnames(df)[2], colnames(descript_data)) # <1>
    descript_data  # <1>
  })  
  output$summary <- renderPrint({
    ddata()
  })
output$downloadSummary <- downloadHandler(  # <2>
    filename = function() {   # <2>
      paste("descriptive-",input$dataset, ".csv", sep = "")  # <2>
    },  # <2>
    content = function(file) {  # <2>
      write.table(ddata(), file, sep = ",",  # <2>
                  row.names = FALSE)  # <2>
    }  # <2>
  )   # <2>
output$downloadPlot <- downloadHandler( # <3>
    filename = function() {  # <3>
      paste0("plot-", input$dataset, ".png") # <3>
    }, # <3> 
    content = function(file) { # <3>
      png(file, width = 800, height = 600) # <3>
      df <- dataInput() # <3>
      plot(df[, 1:2]) # <3>
      dev.off() # <3>
    } # <3>
  ) # <3>
}
```

1. Create a reactive object containing the summary of the data, more suitable for the download
2. Actual download handler, where both the filename and content are defined as functions. The filename includes a brief description and the number of the dataset
3. Same thing, but the dowload handler is designed to handle the download of the plot. The format can be changed 


## Together 
```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```


:::